---
title: "EFA-Boosting Wizard (v2)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{EFA-Boosting Wizard (v2)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

The **EFA-Boosting Wizard** (`run_efa_boosting_wizard()`) is a second Shiny application that coexists with the original `run_efa_boosting()` studio. While the studio is optimised for power users who want a one-screen control panel, the wizard guides the analyst through five sequential phases, each one accompanied by a *"proposed action / what will happen"* explanation. It also adds three capabilities the studio does not have:

1. **Reliability** (categorical omega and Cronbach's alpha by factor) immediately after the boosting step.
2. **Convergent / discriminant validity** with automatic detection of multidimensional comparator instruments and AI-assisted classification.
3. **Downloadable artifacts**: a full TXT audit log of the session and an APA-7 Word manuscript whose tables come from the actual session data (the AI writes only the surrounding prose).

The wizard is fully bilingual (Spanish primary, English) and ships with an **autopilot** mode that drives every step on its own when an OpenAI API key is provided.

## Launching the wizard

```{r}
library(OptimalFactor)
run_efa_boosting_wizard()
```

For headless or programmatic launches (e.g. embedding the app inside Posit Connect or a custom hosting environment) pass `launch.browser = FALSE`:

```{r}
app <- run_efa_boosting_wizard(launch.browser = FALSE)
```

The wizard relies on `shiny`, `bslib` and `DT`; it strongly benefits from `psych`, `lavaan`, `semTools`, `ggplot2` and `commonmark` (chat Markdown rendering). For the AI features (autopilot, conceptual analysis, manuscript drafting) it additionally needs `httr`, `jsonlite`, and for the Word export `officer` plus `flextable`. The autopilot timer uses `later`.

## The five phases

### Phase 1 — Data

Load a CSV / TSV / Excel / RDS file from the sidebar. The wizard auto-detects the items of the scale under analysis by prefix (e.g. typing `EAF` selects `EAF1`, `EAF2`, ...). A manual *override* selector exposes **all** columns of the dataset, mirroring the EasyValidation convention; you can mix and match items that do not share a prefix.

The stats row at the top of the wizard immediately reports the number of items detected, the number of rows, and the active phase.

### Phase 2 — Parallel diagnostic

The wizard runs four independent decision rules in parallel and presents the consensus:

* **Kaiser** — eigenvalues > 1 on the polychoric correlation.
* **Horn's parallel analysis** — eigenvalues compared against a random parallel data set (Horn, 1965).
* **Velicer's MAP** — minimum average partial correlation (Velicer, 1976).
* **BIC** — Schwarz criterion across candidate factor structures.

Each method votes for a number of factors and the wizard recommends the modal choice. The user can accept the recommendation or override it with a theoretical value derived from prior literature.

### Phase 3 — EFA boosting

Runs `efa_boosting()` with the chosen `k`. The verbose stdout produced by the algorithm — including the per-iteration structure, which item was removed and why — is captured into the **Trace** tab. When an OpenAI key is provided, `print_conceptual_analysis()` is appended to the same tab, giving a narrative explanation of why each item was dropped (item-content + statistical reason).

After the boost finishes the user reviews a card with the final fit indices, the number of items retained vs original, and **red pills** listing every removed item. From here the user confirms and moves to reliability.

```{r}
# What happens under the hood (you don't run this yourself in the wizard).
boost <- efa_boosting(
  data        = my_data[, my_items],
  name_items  = "EAF",
  n_factors   = 3L,
  thresholds  = list(loading = 0.30, min_items_per_factor = 3L),
  performance = list(max_candidates_eval = 12L, smart_pruning = TRUE,
                      emit_progress = TRUE),
  verbose     = TRUE)
boost$stop_reason    # canonical reason for stopping
boost$final_structure
```

### Phase 4 — Reliability

For every factor of the boosted solution the wizard reports:

* **McDonald's omega** computed with `semTools::compRelSEM(..., ord.scale = TRUE)`, the categorical variant recommended for ordinal data (Green & Yang, 2009).
* **Cronbach's alpha** from `psych::alpha()` as a complementary indicator.
* The EFA-derived CFA model is fitted with `lavaan::cfa()` so that CFI / TLI / RMSEA / SRMR can be reported alongside the reliability coefficients.

### Phase 5 — Validity

The validity phase is split into two complementary blocks:

**(a) Standalone criterion variables.** A multi-select picker lets the user choose any numeric variable in the dataset (e.g. age, GPA, an external criterion score). The wizard correlates each factor's sum-score with each criterion variable using Pearson's *r*, IC95 % and *p*.

**(b) Comparison instruments (convergent / discriminant).** Click the magnifier icon and the wizard auto-detects candidate instruments from the column names. A hierarchical regex separates **multidimensional** comparators from unidimensional ones:

* `BAI1`, `BAI2`, … `BAI21` → unidimensional instrument `BAI` (21 items).
* `DERS_AC1`, `DERS_OB1`, `DERS_GO1` → multidimensional instrument `DERS` with three subscales (`AC`, `OB`, `GO`).

For each candidate, a quick parallel analysis (Kaiser on the polychoric correlation) is shown as a hint — *"parallel analysis suggests k = 3 (eigen top: 4.2, 1.8, 1.3)"* — so you can confirm or correct the detected dimensionality. When the autopilot is on, OpenAI also classifies each instrument as **convergent** or **discriminant** based on its label (e.g. it knows that BAI measures anxiety, DERS measures emotion regulation).

For every (factor × convergent column) pair the wizard reports:

| Column | Description |
|---|---|
| *r* | Pearson coefficient |
| IC 95 % | Confidence interval |
| *p* | Significance |
| Magnitude | Cohen (1988): *strong* / *moderate* / *small* / *insignificant* |
| Verdict | *Convergence strong / moderate / weak / not supported* (when expected = convergent), or *discriminates / dubious / does NOT discriminate* (when expected = discriminant) |

A heatmap visualises the matrix `principal × convergent`, with asterisks for `p < .05`. The mini-detail is what makes the wizard go beyond a flat sum: if you only correlated the total score of a multidimensional comparator, you'd miss the fact that one of its subscales (e.g. `DERS_GO`) does not relate to your construct even though the rest does.

## Autopilot mode

A checkbox in the sidebar activates the autopilot. When ON, after a configurable read-delay the wizard auto-confirms every step. Three persistent buttons appear on top of every step:

* `← Back` — rewind to the previous phase. Pauses the autopilot for safety. Computed results are kept, not erased.
* `⏸ Pause autopilot` — visible when autopilot is active. Stops the countdown without leaving the phase.
* `▶ Resume AI` — visible when autopilot is paused and AI is still available.

Specifically for Phase 5, the autopilot first calls OpenAI with the detected instruments and a JSON-formatted system prompt asking *"classify each as convergent or discriminant for a scale called X"*. The resulting decisions appear in an orange banner above the configuration card, with each instrument's chip (green for convergent, orange for discriminant) and the AI's textual rationale. The countdown is automatically extended by 4 seconds in this phase to give the user time to review and tweak before the auto-compute fires.

## Downloads

Two large gradient cards live at the top of the **Reporte** panel:

### Session log (`.txt`)

A plain-text audit trail of the entire session: configuration, multi-method consensus details, final loadings (rounded to 2 decimals), per-iteration fit indices, the conceptual AI block (when run), the verbose trace, reliability, external validity, and the convergent / discriminant table with per-pair verdicts. **Does not require an API key.**

### Manuscript (`.docx`)

An APA-7 Word document with two sections drafted by the AI:

* `## 2. Método → 2.4. Análisis de datos` (5–7 dense paragraphs covering software, preprocessing, factor-number consensus, EFA-boosting, reliability and validity).
* `## 3. Resultados` (subsections 3.1 to 3.5; an additional 3.6 *Evidencia convergente y discriminante* is added when comparator instruments are defined).

Tables are inserted via placeholders — `{{TABLE_LOADINGS}}`, `{{TABLE_FIT}}`, `{{TABLE_OMEGA}}`, `{{TABLE_EXTERNAL}}`, `{{TABLE_CONVERGENT}}` — which the runtime replaces with `flextable` objects built from the actual session data. The AI never writes a number on its own; it only writes the surrounding prose. **Requires an OpenAI API key.**

## IA Chat tab

A free-form chat with full context of the fitted model. The system prompt is automatically primed with the dataset shape, the items retained vs removed, the `stop_reason` and the fit indices, so questions like *"why was item EAF6 removed?"* can be answered specifically. Replies are rendered as real HTML via `commonmark::markdown_html()` — `**bold**`, lists, tables, fenced code. Press **Enter** to send; **Shift+Enter** to insert a newline.

## Sample size and execution time

`efa_boosting()` evaluates the top *K* candidate item-removals at each iteration (default `max_candidates_eval = 12`) using smart pruning on the maximum loading. This typically reduces the wall-clock time on long instruments by an order of magnitude. As a benchmark, a 32-item scale that took >30 minutes with the exhaustive search of v1.0 now completes in ~2–3 minutes on the same machine.

The `stop_reason` field of `boost$stop_reason` returned by the boosting reports exactly why the loop ended:

| Value | Meaning |
|---|---|
| `all_criteria_met` | All fit targets satisfied (CFI, TLI, RMSEA, SRMR) |
| `min_items_per_factor_protected` | Cannot remove more without violating the floor |
| `fit_target_reached` | A single fit target was reached and no further gains expected |
| `max_iterations` | Hit the hard-cap on iterations |
| `not_enough_items` | Too few candidate items left |
| `efa_convergence_failed` | EFA model failed to converge |
| `fit_zero_no_structural_problem` | All fit indices are zero, model is saturated |
| `timeout` | Time budget exhausted |

The wizard surfaces this value verbatim in the Phase 3 review card and in the TXT export.

## References

* Cohen, J. (1988). *Statistical power analysis for the behavioral sciences* (2nd ed.).
* Fornell, C., & Larcker, D. F. (1981). Evaluating structural equation models with unobservable variables and measurement error. *Journal of Marketing Research*, 18(1), 39–50.
* Green, S. B., & Yang, Y. (2009). Reliability of summed item scores using structural equation modeling: an alternative to coefficient alpha. *Psychometrika*, 74, 155–167.
* Henseler, J., Ringle, C. M., & Sarstedt, M. (2015). A new criterion for assessing discriminant validity. *Journal of the Academy of Marketing Science*, 43, 115–135.
* Horn, J. L. (1965). A rationale and test for the number of factors in factor analysis. *Psychometrika*, 30, 179–185.
* McDonald, R. P. (1999). *Test theory: A unified treatment*. Erlbaum.
* Velicer, W. F. (1976). Determining the number of components from the matrix of partial correlations. *Psychometrika*, 41(3), 321–327.
